---
title: "Assignment 7 - CME Seat Price Forecast  - Nicholas Petr"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

### Import Required Packages

```{r}
# Import Required packages 
library(forecast)
library(tseries)
library(lubridate)
library(lmtest)
library(TSA)
library(arfima)
library(car)
library(dplyr)
library(ggplot2)
library(zoo)
library(anytime)
library(fGarch)
library(xts)
library(Metrics)
```

### Load Data

```{r}
cmeSP <- read.csv("C:/Users/Nick's Laptop/Desktop/Time Series Analysis/week6_multi-step_forecasting/cmeS.csv")
immSP <- read.csv("C:/Users/Nick's Laptop/Desktop/Time Series Analysis/week6_multi-step_forecasting/immS.csv")
iomSP <- read.csv("C:/Users/Nick's Laptop/Desktop/Time Series Analysis/week6_multi-step_forecasting/iomS.csv")

df_vol <- read.csv("C:/Users/Nick's Laptop/Desktop/Time Series Analysis/week7_intervention_arimaxForecast/Contracts_Volume.csv")
df_Class <- read.csv("C:/Users/Nick's Laptop/Desktop/Time Series Analysis/week7_intervention_arimaxForecast/Contracts_Classification.csv")
```

## You need to write a program, one program should work for all 3 badge types:

### The following program Converts Date field to month/year date, calculates the Floor volume, Aggregates the data for each Commodity Indicator for each month, Sorts out of the volume data, those commodities that are relevant for the particular badge (CME, IMM, IOM), Groups by Date, sum all commodity types for volume 
```{r}
# Convert Date field to month/year date
df_vol$Date = mdy(df_vol$Date)
df_vol$Date = as.Date(as.yearmon(df_vol$Date, "%m/%Y"))

# calculate the Floor volume
df_vol$Electronic.Volume = as.numeric(gsub(",", "", df_vol$Electronic.Volume)) # Convert Electronic volume to numeric
df_vol$Floor.Volume = df_vol$Total.Volume-df_vol$Electronic.Volume

# Aggregate the data for each Commodity Indicator for each month. Don't worry about futures / options, just add them all up.
df = merge(x = df_vol, y = df_Class, by.x = "Commodity.Indicator", by.y = "Commodity.Code", all.x = TRUE)
# Drop unnecessary columns 
df = df[ -c(3:4) ]
# Group by Commodity Indicator, Date, Division and sum the numeric volume columns 
df = df %>% dplyr::group_by(Commodity.Indicator, Date, Division) %>% summarize_all(sum)

# Sort out of the volume data, those commodities that are relevant for the particular badge (CME, IMM, IOM). Keep in mind that the CME can trade EVERYTHING, not just what the list says.
s = split(df, df$Division)
# Separate by Division, drop the division indicator 
# CME
cme = s$CME
cme = cme[ -c(3) ]
# IMM
imm = s$IMM
imm = imm[ -c(3) ]
#IOM
iom = s$IOM
iom = iom[ -c(3) ]

# Groups by Date, sum all commodity types for volume 
cme <-
  cme %>%
  group_by(Date) %>%
  summarise(Electronic.Volume=sum(Electronic.Volume),
            Total.Volume=sum(Total.Volume),
            Floor.Volume=sum(Floor.Volume))
cme

imm <-
  imm %>%
  group_by(Date) %>%
  summarise(Electronic.Volume=sum(Electronic.Volume),
            Total.Volume=sum(Total.Volume),
            Floor.Volume=sum(Floor.Volume))
imm

iom <-
  iom %>%
  group_by(Date) %>%
  summarise(Electronic.Volume=sum(Electronic.Volume),
            Total.Volume=sum(Total.Volume),
            Floor.Volume=sum(Floor.Volume))
iom
```

### Develop Process that interpolates a price for each month for each unique dataset
```{r}
interp = function(df, int_df){
  # manipulation prior to interpolation
  df$DateOfSale = mdy(df$DateOfSale)
  df$date = as.Date(as.yearmon(df$DateOfSale, "%m/%Y"))
  m = seq(df$date[1], tail(df$date,1), by="month")
  
  # interpolation
  df = subset(df, select = c("date", "price"))
  int_df = data.frame(date=m, int_df=spline(df, method="fmm", xout=m)$y)
  return(int_df)
}
cmeSP = interp(cmeSP, int_cme)
immSP = interp(immSP, int_imm)
iomSP = interp(iomSP, int_iom)
```

### Merge interpolated seat prices with volume dataset, Split into Train/Test sets for modeling, Get log values for fGarch. Display CME training set to display output
```{r}
# Merge interpolated seat prices with volume dataset 
cme = merge(x = cme, y = cmeSP, by.x = "Date", by.y = "date")
imm = merge(x = imm, y = immSP, by.x = "Date", by.y = "date")
iom = merge(x = iom, y = iomSP, by.x = "Date", by.y = "date")


# Split into Train/Test sets for modeling 
cme_test <- dplyr::filter(cme, Date >= '2013-01-01')
imm_test <- dplyr::filter(imm, Date >= '2013-01-01')
iom_test <- dplyr::filter(iom, Date >= '2013-01-01')

cme_train <- dplyr::filter(cme, Date < '2013-01-01')
imm_train <- dplyr::filter(imm, Date < '2013-01-01')
iom_train <- dplyr::filter(iom, Date < '2013-01-01')

# Get log values for fGarch
cme_seat_price_xts <- xts(cme_train$int_df, order.by = as.Date(cme_train$Date, "%m/%d/%Y"))
cme_log <- log(cme_seat_price_xts)
cme_log <- diff(cme_log)[-1]

imm_seat_price_xts <- xts(imm_train$int_df, order.by = as.Date(imm_train$Date, "%m/%d/%Y"))
imm_log <- log(imm_seat_price_xts)
imm_log <- diff(imm_log)[-1]

iom_seat_price_xts <- xts(iom_train$int_df, order.by = as.Date(iom_train$Date, "%m/%d/%Y"))
iom_log <- log(iom_seat_price_xts)
iom_log <- diff(iom_log)[-1]

cme_train # Example 
```

## TASK A: Use the trading volume information to forecast the monthly seat price for 2013.

### 1. Linear regression (seat price is dependent, volume(s) independent)
```{r}
fit1_cme <- lm(int_df ~ Total.Volume, data=cme_train) # cme 
summary(fit1_cme)
fit1_imm <- lm(int_df ~ Floor.Volume, data=imm_train) # imm
summary(fit1_imm)
fit1_iom <- lm(int_df ~ Floor.Volume, data=iom_train) # iom
summary(fit1_iom)
```

### 2. Linear regression with ARMA errors (use arima with xreg) 
```{r}
fit2_cme <- auto.arima(cme_train$int_df, xreg=cme_train$Total.Volume)
summary(fit2_cme)
fit2_imm <- auto.arima(imm_train$int_df, xreg=imm_train$Floor.Volume)
summary(fit2_imm)
fit2_iom <- auto.arima(iom_train$int_df, xreg=iom_train$Floor.Volume)
summary(fit2_iom)
```

### 3. Holt Winters
```{r}
fit3_cme <- HoltWinters(ts(cme_train$int_df, frequency = 12))
summary(fit3_cme)
fit3_imm <- HoltWinters(ts(imm_train$int_df, frequency = 12))
summary(fit3_imm)
fit3_iom <- HoltWinters(ts(iom_train$int_df, frequency = 12))
summary(fit3_iom)
```

### 4. ARIMA
```{r}
fit4_cme <- auto.arima(ts(cme_train$int_df, frequency = 12), seasonal = FALSE)
summary(fit4_cme)
fit4_imm <- auto.arima(ts(imm_train$int_df, frequency = 12), seasonal = FALSE)
summary(fit4_imm)
fit4_iom <- auto.arima(ts(iom_train$int_df, frequency = 12), seasonal = FALSE)
summary(fit4_iom)
```

### 5. Seasonal ARIMA (SARIMA) - here seasonality is monthly
```{r}
fit5_cme <- auto.arima(ts(cme_train$int_df, frequency = 12), D=1)
summary(fit5_cme)
fit5_imm <- auto.arima(ts(imm_train$int_df, frequency = 12), D=1)
summary(fit5_imm)
fit5_iom <- auto.arima(ts(iom_train$int_df, frequency = 12), D=1)
summary(fit5_iom)
```

### 6. Fractional ARIMA (ARFIMA) - check applicability first using the ACF
```{r}
acf(ts(cme_train$int_df, frequency = 12))
cme_d <- fracdiff::fracdiff(ts(cme_train$int_df, frequency = 12)) #get the fractional d
cme_d # We can use ARFIMA, as d is .499 (not > .5)
fit6_cme <- forecast::arfima(ts(cme_train$int_df, frequency = 12))
summary(fit6_cme)

acf(ts(imm_train$int_df))
imm_d <- fracdiff::fracdiff(ts(imm_train$int_df, frequency = 12)) #get the fractional d
imm_d # We can use ARFIMA, as d is .499 (not > .5)
fit6_imm <- forecast::arfima(ts(imm_train$int_df, frequency = 12))
summary(fit6_imm)

acf(ts(iom_train$int_df))
iom_d <- fracdiff::fracdiff(ts(iom_train$int_df, frequency = 12)) #get the fractional d
iom_d # We can use ARFIMA, as d is .499 (not > .5)
fit6_iom <- forecast::arfima(ts(iom_train$int_df, frequency = 12))
summary(fit6_iom)
```

### 7. ARMA and GARCH combination - use the fGarch R library and garchFit()
```{r}
# find p,q
cme_arma <- auto.arima(cme_log)
summary(cme_arma)
# input p,q into arma for garch fit
fit7_cme <- garchFit(~arma(0,0) + garch(1,1), data=cme_log, cond.dist = "std", trace=F)
summary(fit7_cme)

# find p,q
imm_arma <- auto.arima(imm_log)
summary(imm_arma)
# input p,q into arma for garch fit
fit7_imm <- garchFit(~arma(0,2) + garch(1,1), data=imm_log, cond.dist = "std", trace=F)
summary(fit7_imm)

# find p,q
iom_arma <- auto.arima(iom_log)
summary(iom_arma)
# input p,q into arma for garch fit
fit7_iom <- garchFit(~arma(0,1) + garch(1,1), data=iom_log, cond.dist = "std", trace=F)
summary(fit7_iom)
```

### Part B: Since you already have the seat prices for the 2013, evaluate each algorithm from Task A using sMAPE. Which one do you recommend to forecast monthly prices for each of the seat classes?

### 1. Linear regression (seat price is dependent, volume(s) independent)
```{r}
cme_fit1_forecast <- predict(fit1_cme, newdata=cme_test)
cme_test$linear_regression <- cme_fit1_forecast
cme_test$linear_regression

imm_fit1_forecast <- predict(fit1_imm, newdata=imm_test)
imm_test$linear_regression <- imm_fit1_forecast
imm_test$linear_regression

iom_fit1_forecast <- predict(fit1_iom, newdata=iom_test)
iom_test$linear_regression <- iom_fit1_forecast
iom_test$linear_regression
```

### 2. Linear regression with ARMA errors (use arima with xreg) 
```{r}
cme_fit2_forecast <- predict(fit2_cme, newxreg=cme_test$Floor.Volume)$pred
cme_test$lm_with_arima_errors <- cme_fit2_forecast
cme_test$lm_with_arima_errors

imm_fit2_forecast <- predict(fit2_imm, newxreg=imm_test$Floor.Volume)$pred
imm_test$lm_with_arima_errors <- imm_fit2_forecast
imm_test$lm_with_arima_errors

iom_fit2_forecast <- predict(fit2_iom, newxreg=iom_test$Floor.Volume)$pred
iom_test$lm_with_arima_errors <- iom_fit2_forecast
iom_test$lm_with_arima_errors
```

### 3. Holt Winters
```{r}
cme_fit3_forecast <- forecast(fit3_cme, h=12)$mean
cme_test$holt_winters <- cme_fit3_forecast
cme_test$holt_winters

imm_fit3_forecast <- forecast(fit3_imm, h=12)$mean
imm_test$holt_winters <- imm_fit3_forecast
imm_test$holt_winters

iom_fit3_forecast <- forecast(fit3_iom, h=12)$mean
iom_test$holt_winters <- iom_fit3_forecast
iom_test$holt_winters
```

### 4. ARIMA
```{r}
cme_fit4_forecast <- forecast(fit4_cme, h=12)$mean
cme_test$arima <- cme_fit4_forecast
cme_test$arima

imm_fit4_forecast <- forecast(fit4_imm, h=12)$mean
imm_test$arima <- imm_fit4_forecast
imm_test$arima

iom_fit4_forecast <- forecast(fit4_iom, h=12)$mean
iom_test$arima <- iom_fit4_forecast
iom_test$arima
```

### 5. Seasonal ARIMA (SARIMA) - here seasonality is monthly
```{r}
cme_fit5_forecast <- forecast(fit5_cme, h=12)$mean
cme_test$sarima <- cme_fit5_forecast
cme_test$sarima

imm_fit5_forecast <- forecast(fit5_imm, h=12)$mean
imm_test$sarima <- imm_fit5_forecast
imm_test$sarima

iom_fit5_forecast <- forecast(fit5_iom, h=12)$mean
iom_test$sarima <- iom_fit5_forecast
iom_test$sarima
```

### 6. Fractional ARIMA (ARFIMA) - check applicability first using the ACF
```{r}
cme_fit6_forecast <- forecast(fit6_cme, h=12)$mean
cme_test$arfima <- cme_fit6_forecast
cme_test$arfima

imm_fit6_forecast <- forecast(fit6_imm, h=12)$mean
imm_test$arfima <- imm_fit6_forecast
imm_test$arfima

iom_fit6_forecast <- forecast(fit6_iom, h=12)$mean
iom_test$arfima <- iom_fit6_forecast
iom_test$arfima
```

### 7. ARMA and GARCH combination - use the fGarch R library and garchFit()
```{r}
cme_fit7_forecast <- predict(fit7_cme, n.ahead=12)$meanForecast
cme_test$garch <- cme_fit7_forecast
cme_test$garch

imm_fit7_forecast <- predict(fit7_imm, n.ahead=12)$meanForecast
imm_test$garch <- imm_fit7_forecast
imm_test$garch

iom_fit7_forecast <- predict(fit7_iom, n.ahead=12)$meanForecast
iom_test$garch <- iom_fit7_forecast
iom_test$garch
```

### Calculate modeling outputs using sMAPE
```{r}
smape_results <- data.frame(
  row.names = c('CME', 'IMM', 'IOM'),
  linear_regression = c(smape(iom_test$Floor.Volume, iom_test$linear_regression),
                        smape(imm_test$Floor.Volume, imm_test$linear_regression), 
                        smape(cme_test$Floor.Volume, cme_test$linear_regression)),
  linear_regression_with_arima_errors = c(smape(iom_test$Floor.Volume, iom_test$lm_with_arima_errors), 
                                          smape(imm_test$Floor.Volume, imm_test$lm_with_arima_errors), 
                                          smape(cme_test$Floor.Volume,        cme_test$lm_with_arima_errors)), 
  holt_winters = c(smape(iom_test$Floor.Volume, iom_test$holt_winters),
                   smape(imm_test$Floor.Volume, imm_test$holt_winters), 
                   smape(cme_test$Floor.Volume, cme_test$holt_winters)),
  arima = c(smape(iom_test$Floor.Volume, iom_test$arima),
            smape(imm_test$Floor.Volume, imm_test$arima), 
            smape(cme_test$Floor.Volume, cme_test$arima)),
  sarima = c(smape(iom_test$Floor.Volume, iom_test$sarima),
             smape(imm_test$Floor.Volume, imm_test$sarima), 
             smape(cme_test$Floor.Volume, cme_test$sarima)),
  arfima = c(smape(iom_test$Floor.Volume, iom_test$arfima),
             smape(imm_test$Floor.Volume, imm_test$arfima), 
             smape(cme_test$Floor.Volume, cme_test$arfima)),
  garch = c(smape(iom_test$Floor.Volume, iom_test$garch),
             smape(imm_test$Floor.Volume, imm_test$garch), 
             smape(cme_test$Floor.Volume, cme_test$garch))
) 


smape_results
```

### Which one do you recommend to forecast monthly prices for each of the seat classes?

Looking at the table outputs you would use ARFIMA for CME seat prices, ARFIMA for IMM prices, and SARIMA for IOM prices.